---
title: "Practicing with ggplot2"
output: html_document
date: "2023-06-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We're continuing following along with The Book of R, but now we're getting familiar
with the ggplot2 package.  There are distinct advantages to using this package to 
create and manipulate graphics as compared to the base R graphics.  The main distinction 
being that ggplot2 stores plots as objects.

First, lets get our environment set up by loading the packages we'll need as well
as storing the data we'll be plotting.
```{r}
library("tidyverse")
```
```{r}
foo <- c(1:9)
bar <- c(6, 3, 5, 10, 4, 8, 2, 9, 3)
fbar <- cbind(foo, bar)
```

Perfect, now let's see the basic ggplot2 plot fxn 'qplot' in action.
```{r}
qplot(foo, bar)
```

Okay, cool. We have a nice collection of points. We can also assign this plot to an object, 
and add some labels in the same way we did with R's basic 'plot' fxn.
```{r}
q <- qplot(foo, bar, main= "My qplot", xlab= "x-axis", ylab= "y axis")
q
```

Just like the book said, the plot didn't automatically populate when I assigned it to 'q',
I had to 'print' q to display it.

Now, let's spice it up with some features using the 'geom_point()' and 'geom_line()' 
fxns. The first chunck of code with print the default settings for both of these fxns.
Then we'll specify different styles in a second code chunck.
```{r}
qplot(foo, bar, geom= "blank") + geom_point() + geom_line()
```
```{r}
qplot(foo, bar, geom= "blank") + geom_point(size= 3, shape= 6, color= "blue") +
  geom_line(color= "red", linetype= 2)
```

The object-oriented nature of ggplot2 allows you to store a plot and then tweak one the one 
section of code you want to play around with. For example, if we wanted to adjust the 
parameters of the line part of the above graph, we can store the main qplot fxn, along 
with the geom_point() parameters, then tweak the line component of code.
```{r}
q2 <- qplot(foo, bar, geom= "blank") + geom_point(size= 3, shape= 6, color= "blue")
q2 + geom_line(color= "seagreen4", linetype= 1)
```
```{r}
q2 + geom_line(color= "seagreen4", linetype= 3)
```

## Complex Graphs
Okay, now we're going to get into more complex plots, by essentially mapping the aesthetic 
parameters to a dataset that is split or grouped, like a factor object. Let's enter 
in the same data we did for the last plotting practice to construct that elaborate plot.
```{r}
x <- 1:20
y <- c(-1.49, 3.37, 2.59, -2.78, -3.94, -0.92, 6.43, 8.51, 3.41, -8.23, -12.01, -6.58,
       2.87, 14.12, 9.63, -4.58, -14.78, -11.67, 1.17, 15.62)
```

Now that we have our data entered, we can explicitly define the factor as follows:
```{r}
ptype <- rep(NA, length(x=x))
ptype[y>=5] <- "too big"
ptype[y<=-5] <- "too small"
ptype[(x>=5 & x<=15) & (y>-5 & y<5)] <- "sweet"
ptype[(x<5 | x>15) & (y>-5 & y<5)] <- "standard"
ptype <- factor(x=ptype)
ptype
```

Cool, we now have a factor of 20 observations, sorted into 4 levels, we can now plot a simple 
graphic with qplot.
```{r}
qplot(x, y, color= ptype, shape= ptype)
```

That's pretty cool, and somewhat simpler than using the basic R 'plot()' fxn. Crafting the 
factor seemed involved, but it was a good exercise for subsetting.  

Now, let's add to this graph to make it look like complex graph we created in the 
previous practice using 'plot()'.
```{r}
qplot(x, y, color= ptype, shape= ptype) + geom_point(size=4) + 
  geom_line(mapping= aes(group=1), color= "black", lty= 2) +
  geom_hline(mapping= aes(yintercept= c(-5, 5)), color= "red") +
  geom_segment(mapping= aes(x= 5, y= -5, xend= 5, yend= 5), color= "red", lty= 3) +
  geom_segment(mapping= aes(x= 15, y= -5, xend= 15, yend= 5), color= "red", lty= 3)
```

Wow, beautiful! I MUCH prefer this graph to the one manually crafted using the 'plot()' fxn.