---
title: 'Intro to R: Data Visualization Lab'
author: "Hannah Bravo De Rueda"
date: "2023-09-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Creating Visualizations using ggplot2

#### Question 1: 
 Let's recreate a plot using the mpg dataset.  
 
 The plot is a scatterplot with a trendline going through the points.  'Displ' is mapped to the x-axis, and 'hwy' to the y-axis. Let's give it a try!
 
```{r scatterplot}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() + geom_smooth(se = FALSE)

```
 
 Yayy!  Came out perfect.
 
 #### Question 2: 
 
 We're going to work off the last plot, but now map a third variable 'drv' to color. 
```{r color}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(aes(color = drv)) + geom_smooth(se = FALSE)
```
 
Awesome! I had to make sure to add the color specification to ONLY the geom_plot section. 

#### Question 3: 

Next, we're going to continue building off the previous graph, but now we're going to differentiate the trendline across 'drv' as well. 
```{r linetype}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(aes(color = drv)) + geom_smooth(aes(linetype = drv), se = FALSE)
```

Cool!  So far, everything is working!

#### Question 4: 

We're going to continue working with the mpg dataset, but we're going to move onto a new visualization now.  

This new plot is still a scatterplot, with 'cty' mapped to the x-axis, 'hwy' mapped to the y-axis, and 'class' mapped to color. Let's give it a try!
```{r jitterplot}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy, color = class)) +
  geom_point(position = "jitter")
```

Awesome!  I originally plotted it as a scatterplot, but since both variables are not continuous, so it graphed various points over one another and didn't show ALL of the data points.  Using geom_jitter(), kind of spaced some out the points out from one another, showing much more of the data. Looking back, you could change geom_point() to geom_jitter(), OR you could add the argument 'position = "jitter"' to geom_point().  They both produce the same plot.

#### Question 5: 
 Okay, now we're going to distinguish each class as it's own facet, rather than using color, and we're going to flip the coordinates.  Let's see what we get!
```{r facet_wrap}
ggplot(data = mpg, mapping = aes(x = hwy, y = cyl)) +
  geom_point(color = "orange", shape = 17, show.legend = FALSE) + facet_wrap(~class) 
```
 
 Yayy!! That took a bit longer, but we got there.  I was originally including the color and shape parameters in the 'aes' which was a mistake. 